Modeling industrial symbiotic processes from a complex systems perspective

### 2016-12-30 paper.tex
Update on Overleaf.
.

\title{Complexity Insights into Circular Economy}

\author[1]{Joris Broere}
\author[2]{Christine Moore}
\author[3]{Juste Raimbault}
\author[4]{Jesus Mario Serna}
\author[5]{Marius Somveille}
\author[6]{Evelyn Strombom}
\author[7]{Lorraine Sugar}
\author[8]{Ben Zhu}

\affil[1]{Affiliation, department, city, postcode, country}
\affil[2]{Affiliation, department, city, postcode, country}

@@

\affil[+]{these authors contributed equally to this work}

@@

\begin{abstract}
The 'circular economy' is recent approach on questions of sustainability. The current 'take, make, waste' way of producing is considered to be not sustainable anymore. Current thoughts on 'closing the loop' mechanisms on producing and consuming are booming in many different academic fields. However, not yet from an integrated point of view. The idea of 'closing the loop' mechanisms corresponds to complex system paradigms, however complexity science has hardly been utilized in this field. In this project we started brainstorming on how complexity science can contribute to the field of circular economy. We think there are many ways in which complexity science can contribute, both in methods and in theory.

In the past four weeks at CSSS 2016 we focused by modeling the circular economy by means of an agent based model. So far we have made an toy model that models waste flow between actors on the level of organizations/businesses. The purpose of this model is to investigate how waste flow is affected by different parameters such as, distance, clustering, transportation cost and other parameters. After the CSSS 2016 we will continue to work on the model, it will be validated by applying it to real world spatial maps and infrastructures, and calibrate it to data of first implementations of `circular economy practices'. The goal is to make a basis for an open source circular economy application that can be used to monitor the circular economy, as well as create a market place for waste products. In what follows a first draft of the paper of the toy model.

\end{abstract}

@@

The notion of a circular economy proposes that traditionally separate industries act collectively to obtain a mutual competitive advantage through their physical exchanges of materials, energy, water, and byproducts~\cite{chertow2000industrial}. Described as a material symbiosis, a circular economy approach seeks to recycle residual waste or by-products through the development of complex interlinkages between companies and firms.  In direct contrast with the conventional linear economic approach of material production of produce-use-dispose, the circular economy concept seeks to reduce the uptake of virgin materials while also reducing total waste production by accommodating the cycle of materials (EU Document). This concept is distinct from traditional recycling whereby products are often reduced to their lowest nutrient level, and then disposed of. In a circular economy approach, waste or by product materials of one firm have the potential to become high nutrient level inputs to another firm. As a result, there isn’t a sequential downgrading of waste or by-products, but rather a full cycling of materials.

%%%%%%%%%%%%%%%%%%%%%
% Prospective Literature review 
%%%%%%%%%%%

% could we copy-paste here the literature review that was done in the google doc ?-> would give a bit more background

%\textit{This section gives a broad prospective literature review, not necessarily structured.}

\subsubsection*{Existing Works}

\paragraph{Urban Metabolism}

@@

\cite{olsen1982urban} give theory of the city as an organism, associated temporal and spatial scales, corresponding cycles, economic variables and human scales. Cycles are detailed, but no numerical application nor conceptual model, stays at the qualitative descriptive level. Can be used in our case as a “theoretical backup” reference.

@@

\cite{bai2007industrial} is an editorial for journal of Indus. Ecology (close to an opinion paper). Relatively close to our position, as argues that study of cities and industrial ecology have to be reconciled ($\simeq$ look at industrial ecology from a complexity perspective). ``Tackling global problems at the local level.'' is in the spirit of our rationale.

@@

\cite{li2015resilience} : this paper does simulations based on real data on the resilience of industrial symbioses networks by looking at what happens if a node get removed.

@@

This approach has been practically adopted on a small scale in Asia, specifically in China, via the establishment of eco-industrial parks.

@@

@@

@@

%%%%%%%%%%%%%%%%%%
\subsubsection*{Model setup}

@@

%%%%%%%%%%%%%%%%%%
\subsubsection*{Growing the circular economy network}

@@

The utility function associated with each potential exchange of by-products between two companies was defined as follows:
\[u = \frac{o - c*d}{d_{max}}\]
where $o$ is the overlap between the two companies in by-product space, $c$ is the transportation cost, $d$ is the geographical distance between the two companies, and $d_{max}$ is the maximum distance between any two companies in the system.

In our model, at each time step, the following set of rules was applied:
\begin{itemize}
\item A company, the ``current contractor'' is drawn at random % distrib
\item Potential partners are drawn
\begin{itemize}
\item according to geographical interaction potential $V_{ij}=\frac{1}{d_{ij}^{\alpha}}$
\item among these, the ones whose overlap is above $T_o$
% detail the overlap in terms of conditional probabilities
% and update of "remaining distributions"
\end{itemize}
\item given the set of utilities $(u_{1j},u_{j1})_j \simeq (u_j)_j$, the effective partner is chosen with best utility
% - note : write on potential bargain game
% - write utility with transportation cost
% - explain why utility equal
% - transportation cost are shared : LOGICAL because one can pick up, or the other deliver == shared
% RANDOM UTILITY MODEL
\end{itemize}

\subsubsection*{Indicators of Circularity}

@@

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{figures/Variables.png}
\caption{\label{fig:Variables}Variables used in calculation of indicators, as determined for one trade between companies (redraw for final paper). Total material input, $IM$, is the sum of input material for all companies; total recycled materials, $RM$, is the sum of recycled material for all companies; the total waste, $W$, is the sum of waste for all companies.}
\end{figure}

In their study, \cite{haas2015circular} estimate these indicators for the global and European economies. The authors found that the total processed materials in the global economy is 62 Gt/year (58 Gt/year raw material plus 4 Gt/year recycled), and the degree of circularity is 6 percent. The European Union was found to have 7.7 Gt/year of total processed materials with a degree of circularity of 13 percent. Waste throughput for both economies was found to be 66 percent. These real-economy values provide a point of comparison for the model results.

\paragraph{Geographical setup}

@@

A first prototype was implemented in \texttt{NetLogo}. Further developments will include a \texttt{R} version, especially for the integration into a \texttt{shiny} web application for a real world use as described before. Model exploration was done using model exploration software OpenMole~\cite{reuillon2013openmole}. Code and results are available on the open repository of the project\footnote{at \texttt{https://github.com/SFICSSS16-CircularEconomy/CircularEconomy}}.

@@

%%%%%%%%%%%%%%%%%%
\subsubsection*{Internal model validation}

@@

@@

%%%%%%%%%%%%%%%%%%
\subsection*{Possible Extensions}

@@

@@

>>>>

% writing utilities

% comments and responses
%  -> use this comment to ask questions on what other wrote/answer questions with optional arguments (up to 4 answers)
\usepackage{xparse}
\DeclareDocumentCommand{\comment}{m o o o o}
{%
    \textcolor{red}{\textbf{C : }#1}
    \IfValueT{#2}{\textcolor{blue}{\textbf{A1 : }#2}}
    \IfValueT{#3}{\textcolor{ForestGreen}{\textbf{A2 : }#3}}
    \IfValueT{#4}{\textcolor{red!50!blue}{\textbf{A3 : }#4}}
    \IfValueT{#5}{\textcolor{Aquamarine}{\textbf{A4 : }#5}}
}

% todo
\newcommand{\todo}[1]{\textcolor{red!50!blue}{\textbf{TODO : \textit{#1}}}}

%%%%%%%%%%%%%%%%%%%%%

\title{%Working paper: 
Modeling industrial symbiotic processes from a complex systems perspective}

\author[1,*]{Joris Broere}
\author[2]{Christine Moore}
\author[3,4]{Juste Raimbault}
\author[5]{Jesus Mario Serna}
\author[6]{Marius Somveille}
\author[7]{Evelyn Strombom}
\author[8]{Lorraine Sugar}
\author[9]{Ben Zhu}

\affil[1]{Affiliation, department, city, postcode, country}
\affil[2]{Affiliation, department, city, postcode, country}
\affil[3]{UMR CNRS 8504 G{\'e}ographie-cit{\'e}s, Paris, France}
\affil[4]{UMR-T 9403 IFSTTAR LVMT, Champs-sur-Marne, France}
\affil[5]{Affiliation, department, city, postcode, country}
\affil[6]{Affiliation, department, city, postcode, country}
\affil[7]{Affiliation, department, city, postcode, country}
\affil[8]{Affiliation, department, city, postcode, country}
\affil[9]{Affiliation, department, city, postcode, country}

@@

%\affil[+]{these authors contributed equally to this work}

@@

\begin{abstract}
The 'circular economy' is recent approach on questions of sustainability. The current 'take, make, waste' way of producing is considered to be not sustainable anymore. Current thoughts on 'closing the loop' mechanisms on producing and consuming are booming in many different academic fields. However, not yet from an integrated point of view. The idea of these mechanisms corresponds to complex system paradigms, however complexity science has hardly been utilized in this field.
%In this project we started brainstorming on how complexity science can contribute to the field of circular economy. We think there are many ways in which complexity science can contribute, both in methods and in theory. 
%In the past four weeks at CSSS 2016 we focused by modeling the circular economy by means of an agent based model.
We describe a toy model to model waste flow between actors on the level of organizations/businesses. Its purpose is to investigate how waste flow is affected by different parameters such as, distance, clustering, transportation cost and other parameters.
%After the CSSS 2016 we will continue to work on the model,
Future developments will include the application to real world spatial maps and infrastructures, and calibration to data of first real-world implementations of `circular economy practices'. The goal is to make a basis for an open source circular economy application that can be used to monitor the circular economy, as well as create a market place for waste products.
%In what follows a first draft of the paper of the toy model.
\end{abstract}

@@

The notion of a circular economy proposes that traditionally separate industries act collectively to obtain a mutual competitive advantage through their physical exchanges of materials, energy, water, and byproducts~\cite{chertow2000industrial}. Described as a material symbiosis, a circular economy approach seeks to recycle residual waste or by-products through the development of complex interlinkages between companies and firms.  In direct contrast with the conventional linear economic approach of material production of produce-use-dispose, the circular economy concept seeks to reduce the uptake of virgin materials while also reducing total waste production by accommodating the cycle of materials (EU Document).\comment{need the ref}
 This concept is distinct from traditional recycling whereby products are often reduced to their lowest nutrient level, and then disposed of. In a circular economy approach, waste or by product materials of one firm have the potential to become high nutrient level inputs to another firm. As a result, there isn’t a sequential downgrading of waste or by-products, but rather a full cycling of materials.

%%%%%%%%%%%%%%%%%%%%%
% Interdisicplinary Literature review 
%%%%%%%%%%%

\subsubsection*{Existing Approaches}

An interdisciplinary review on Circular Economy is done in~\cite{ghisellini2016review}.

\paragraph*{Urban Metabolism}

@@

Olsen develops in \cite{olsen1982urban} a theory of the city as an organism. The metaphor yields associated temporal and spatial scales, corresponding cycles for economic and human variables. Cycle contents are qualitatively detailed, but no numerical application nor conceptual model are provided. Can be used in our case as a “theoretical backup” reference.

@@

The field of industrial ecology begins to self-witness its own limitations and the need for more opening and integrative methods. Indeed, in \cite{bai2007industrial}, as an editorial for the Journal of Industrial Ecology, it is argued that the study of cities and industrial ecology have to be reconciled. The proposition of ``Tackling global problems at the local level'' corresponds to our idea of a complexity-rooted bottom-up approach to circular economy.

@@

The resilience of industrial symbioses networks is studied by~\cite{li2015resilience}, through simulations based on real data by looking at what the impact of node removal.

@@

\paragraph{Real-world Applications}

This approach has been practically adopted on a small scale in Asia, specifically in China, via the establishment of eco-industrial parks~\cite{geng2013measuring}.

@@

\cite{george2015circular} : model of economic growth with circular economy

We propose thus in this monograph to tackle this issue at a modest level, by exploring patterns of theoretical feasibility from an agent-based modeling point of view.

It has been recently postulated that evidence-based methods, in particular agent-based modeling, could be crucial for economy in general~\cite{farmer2009economy}.

@@

%%%%%%%%%%%%%%%%%%
\subsection*{Rationale}

% develop the basic idea behind our model ; and why we think results from agent-based simulations can give useful insight into dynamics of circular economy and its feasibility.

@@

%%%%%%%%%%%%%%%%%%
\paragraph*{Model setup}

@@

%%%%%%%%%%%%%%%%%%
\paragraph*{Growing the circular economy network}

@@

The utility function associated with each potential exchange of by-products between two companies was defined as follows:
\[
u = o - c \cdot \frac{d}{d_{max}}
\]
where $o$ is the overlap between the two companies in by-product space, $c$ is the transportation cost, $d$ is the geographical distance between the two companies, and $d_{max}$ is the maximum distance between any two companies in the system.

In our model, at each time step, the following set of rules was applied:
\begin{itemize}
\item A company, the ``current contractor'' is drawn at random % distrib
\item Potential partners are drawn
\begin{itemize}
\item according to geographical interaction potential $V_{ij}=\frac{1}{d_{ij}^{\alpha}}$
\item among these, the ones whose overlap is above $T_o$ are taken as potential partners
% detail the overlap in terms of conditional probabilities
% and update of "remaining distributions"
\end{itemize}
\item given the set of utilities $(u_{1j},u_{j1})_j \simeq (u_j)_j$, the potential partner with best utility is chosen
% - note : write on potential bargain game
% - write utility with transportation cost
% - explain why utility equal
% - transportation cost are shared : LOGICAL because one can pick up, or the other deliver == shared
% RANDOM UTILITY MODEL
\end{itemize}

\paragraph*{Indicators of Circularity}

@@

%%%%%%%%%%%%%%%%%
\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{figures/Variables.png}
\caption{\label{fig:Variables}Variables used in calculation of indicators, as determined for one trade between companies (redraw for final paper). Total material input, $IM$, is the sum of input material for all companies; total recycled materials, $RM$, is the sum of recycled material for all companies; the total waste, $W$, is the sum of waste for all companies.}
\end{figure}
%%%%%%%%%%%%%%%%%

In their study, \cite{haas2015circular} estimate these indicators for the global and European economies. The authors found that the total processed materials in the global economy is 62 Gt/year (58 Gt/year raw material plus 4 Gt/year recycled), and the degree of circularity is 6 percent. The European Union was found to have 7.7 Gt/year of total processed materials with a degree of circularity of 13 percent. Waste throughput for both economies was found to be 66 percent. These real-economy values could provide a point of comparison for further development of the model that would include real world economic data.

\paragraph*{Geographical setup}

@@

A first prototype was implemented in \texttt{NetLogo}. A \texttt{R} version was also developed, especially with the objective of an integration into a \texttt{shiny} web application for a real world use as described before. Model exploration was done using model exploration software OpenMole~\cite{reuillon2013openmole}. Code and results are available on the open repository of the project\footnote{at \texttt{https://github.com/SFICSSS16-CircularEconomy/CircularEconomy}}.

%%%%%%%%%%%%%%%%%%
\subsection*{Internal model validation}

\paragraph*{Statistical Consistency}

\paragraph*{Path-dependency and Emergence}

% why abm is necessary

@@

@@

%%%%%%%%%%%%%%%%%%
\subsection*{Patterns of Policy Optimization to Grow the Circular Economy}

@@

%%%%%%%%%%%%%%%%%%
\subsection*{Model Extensions}

@@

%%%%%%%%%%%%%%%%%%
\subsection*{Model Application}

% directions towards an application on real data, what scale, etc.

%%%%%%%%%%%%%%%%
\section*{Conclusion}

@@

%%%%%
%%  Supplementary Materials

\newpage

%%%%%%%%%%%%%%%%
\section*{Supplementary Materials}
